From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-19 Pi+/paper/finite.tex
Sorting examples

Because $\Pi$ The group that we are interested in is the group of permutations on a
fininte set. We can define it formally in two different ways:
8a. Taking the type Aut (Fin n) and showing that it satisfies the group
axiom.
8b. Writing up a presentation - choosing generating set and equations that
have to be satisfied.

9. By proving these two to be equivalent, we are bridging the gap between
syntactic and semantic side in the whole proof - by assinging the meaning
(a bijection) to a word (a program).

10. However, in the point 8b, there is an element of choice. One group can
be presented in many different ways. For example, we could generate the
permutations by using generators that:
 - swap i-th element with j-th element
 - swap i-th element with the first element
 - shift elements in a cyclic way: operation with label "i" leaves elements
with numbers > i in place, and element j <= i goes to (j + 1) mod i.

11. One way of thinking about these presentations is to compare them to
sorting algorithms with different primitive operations. If using a set of
operations, one can sort any list, then these operations generate the
permutations group. One weird example: pancake sorting.

>>>>

Because $\PiLang$ is used for reasoning about reversible functions on finite
sets, the group that we are interested in is the group of permutations on a
fininte set, which is classically know as the symmetric group on $n$ words,
$\Sn$. We've already seen that we can define it formally by taking the
automorphism group given by the type $\Aut[\Fin[n]]$, and showing that it has a
group structure satisfying the group axioms. Alternatively, we can write a group
presentation for $\Sn$ -- by defining a set of generators and a set of relations.

The technical contribution of this section is a proof that these two
descriptions are equivalent. By doing that, we bridge the gap between the
syntactic and semantic notions in our completeness proof -- by establishing a
correspondence between the meaning of a program (a bijection) and its syntax (a
word).

However, in giving a group presentation, there is an element of choice. A group
can be presented in many different ways. For example, we could generate the
permutation group on $\Fin[n]$ by using generators that:

\begin{itemize}
\item swap the $i$-the element with the $(i+1)$-th element, that is, adjacent swaps, or
\item swap the $i$-th element with the $j$-th element, for arbitrary $i$s and $j$s, or
\item swap the $i$-th element with an element at a fixed position, or
\item flip a prefix $\Fin[k]$ of $\Fin[n]$ for $k \leq n$, or
\item cyclically shift any subset of $\Fin[n]$.
\end{itemize}

One way of thinking about these presentations is via sorting algorithms, which
use different primitive operations. A sorting algorithm has to
calculate a permutation of a list or a finite set, which satisfies the invariant
of being a sorted sequence, which means, the primitive operations of a sorting
algorithm should be able to generate all the permutations on a given list. If
using a chosen set of reversible operations, one can write a sorting algorithm,
then those operations generate the permutation group.

For example, bubble sort uses the primitive operation of adjacent swaps,
insertion sort and selection sort use the primitive operation of swapping the
$i$-th element with the $j$-th element, cycle sort uses cyclical shifts of
subsequences, pancake sort uses flips of prefixes of the list, et cetera.
\todo{check!}
